Dispersion Distance and the Matter Distribution of the Universe in Dispersion Space

### 2015-06-04 writeup/dm_space_distortions.tex
dispersion measure -> dispersion

However, redshift is not the only measure of radial distance.   Standard
candles and standard rulers can be used to estimate distance using brightness
and angular size respectively.  In the same way a \emph{standard ping}, a brief
broadband radio impulse, can be used to estimate radial distance because of the
characteristic $\lambda^2$ dispersion of electromagnetic signals along the
line-of-sight as a function of wavelength.  This technique, routinely done with pulsars in the Galaxy, can be used to define
a \emph{dispersion distance} to cosmological sources.  Intriguingly, the dispersion of fast radio bursts (FRB), a new class
of millisecond duration radio transients first reported by \citet{2007Sci...318..777L}, has been measured with high significance. 
These bursts are the archetype of a standard ping with 
10 reported
\citep{2012MNRAS.425L..71K,2013Sci...341...53T,
2014ApJ...790..101S, 2014ApJ...792...19B, 
2015MNRAS.447..246P,2015ApJ...799L...5R} to date and total rate of bursts
been estimated to be $3.3^{+5}_{-2.5} \times 10^3$
per sky per day \cite{2015arXiv150500834R}. 
Dispersion is characterized by a line-of-site integral of the free electron density known the dispersion measure
(DM).
FRB events are observed to have dispersion measures of order
$\sim\!\!1000 {\rm pc} / {\rm cm}^3$, greatly in excess of
the galactic expectation of order $\sim\!\!100\,{\rm pc} / {\rm cm}^3$ (depending on
galactic latitude)
\cite{2014arXiv1412.4829D}.
They  thus appear to be of extra-galactic origin, with their
dispersion produced either by the
inter-galactic medium (IGM) \citep{2013Sci...341...53T} surrounding galaxies or the
environment of the source
\citep{2014ApJ...785L..26L,2015arXiv150101341P,2015arXiv150505535C}.
If the dispersion is dominated by electrons along the line-of-sight,
then the average DM--distance relation can be modelled
\citep{2014ApJ...797...71Z,2004MNRAS.348..999I, 2003ApJ...598L..79I}
and the sources appear to be at
cosmological distances of of order $\sim\!\!{\rm Gpc}$. Such cosmological distances permit
the study of clustering of FRB sources in three dimensions with an expectation that these events trace the
large-scale structure of the Universe on linear scales. Note that although we
frame our discussion in terms of FRBs, other bright radio transients that fluctuate on timescales less than the typical lag time of might also be used as standard pings.

Neither redshift nor dispersion measure are perfect proxies for radial distance.
Redshifts are systematically biased by the peculiar velocities of the object
relative to the Hubble flow. This leads to additional apparent clustering
in redshift
space which was first calculated by \citet{1987MNRAS.227....1K}
using linear theory.
Similarly, distance estimates from dispersion measure will be biased by
inhomogeneities in the electron density \citep{2014ApJ...780L..33M},
again leading to additional apparent
clustering of sources in dispersion measure space.  In this letter we calculate
these DM-space distortions and consider the detectability of the signal by
upcoming surveys.

%\section{Density perturbations in dispersion measure space}
\section{Clustering in dispersion measure space}

@@

\emph{Dispersion measure space} is the three dimensional coordinates,
$\vec x_s$ inferred from the dispersion measure assuming that the electrons are
homogeneous. This only affects the radial coordinate such that $\vec x_s = \hat
n \chi_s$ with $\chi_s$ defined by the equation
\begin{equation}
    {\rm DM}(\chi_s) = \int_0^{\chi_s} \ud\chi' a(\chi')^2 \bar{n}_e(\chi').
\end{equation}

@@

We wish to relate the density of a tracer, $f$, measured in DM space to its
density in real space. We follow the derivation in \citet{1987MNRAS.227....1K}
of the
redshift-space distortions. Start by noting that the total number of tracers
in a volume element is the same in both spaces:
\begin{equation}
\label{e:density}
n_{fs}(\vec x_s) \ud^3\vec x_s = n_{f}(\vec x) \ud^3\vec x.
\end{equation}
We split the density into a homogeneous part plus perturbations,
\begin{equation}
    \bar{n}_{fs}(\chi_s)\left[ 1 + \delta_{fs}(\vec x_s)\right] \ud^3\vec x_s
    = \bar{n}_{f}(\chi)\left[ 1 + \delta_f(\vec x)\right] \ud^3\vec x.
\end{equation}
Averaged over the sky, $\langle \chi_s \rangle = \chi$ and thus the background
density should
be the same in DM space as in real space,
\begin{equation}
\bar{n}_{fs}(\chi) = \bar{n}_f(\chi).
\end{equation}
Therefore,
\begin{align}
\bar{n}_{fs}(\chi_s) 
    &= \bar{n}_{f}(\chi) + (\chi_s - \chi)\frac{\ud \bar{n}_f}{\ud \chi}\\
    &= \bar{n}_{f}(\chi)
       + \frac{\ud \bar{n}_f}{\ud \chi}\int_0^\chi \ud \chi' \delta_e(\hat n
       \chi').
       \label{e:nfs}
\end{align}

@@

Substituting Equations~\ref{e:nfs}~and~\ref{e:jac} into Equation~\ref{e:density},
we obtain
\begin{equation}
\label{e:delta_s}
    \delta_{fs} = \delta_f - \delta_e
    - \left(\frac{1}{\bar{n}_f}\frac{\ud \bar{n}_f}{\ud \chi}
    + \frac{2}{\chi} \right)
        \int_0^\chi \ud \chi' \delta_e(\hat n \chi').
\end{equation}
In the equation above, the $-\delta_e$ term is most analogous to the Kaiser
redshift-space distortions.  It is a dilution of tracers in DM space due to an
excess of electrons between the tracers. However we note that this term is isotropic
in contrast to Kaiser.  This is because any wave vector electron perturbation
causes DM space distortions, whereas the radial velocities that cause
redshift-space distortions are only sourced by perturbations with radial wave
vectors.

@@

Large-scale structure is usually studied through its two-point statistics, most
commonly the power spectrum, $P(k)$.
Unlike in redshift-space distortions, Equation~\ref{e:delta_s} does not have a
simple form in harmonic space. The equation's third term couples harmonic modes
and thus the two-point statistics cannot be phrased as a simple power spectrum.
We will instead use $C^{ss}_\ell(\chi,\chi')$, which is
the cross-correlation angular power spectrum of the DM-space over density,
on shells at $\chi$ and $\chi'$:
\begin{align}
    &C^{ss}_\ell(\chi, \chi') =
    \int\ud\Omega\ud\Omega' Y_{\ell m}(\hat n) Y^*_{\ell m}(\hat n')
    \langle \delta_{fs}(\hat n \chi) \delta_{fs}(\hat n' \chi')
        \rangle.
\end{align}
The first two terms in Equation~\ref{e:delta_s} are stationary. For stationary,
isotropic, tracers $x$ and $y$ we have 
$\langle \delta_x(\vec k, \chi) \delta_y(\vec k', \chi) \rangle = (2\pi)^3
\delta^3(\vec k - \vec k') P_{xy}(k, \chi)$.  If for a moment we ignore time
dependence, the angular cross-power spectrum of such tracers is
\begin{align}
    %\delta_{\ell\ell'}\delta_{mm'}C^{xy}_\ell(\chi, \chi')
    %&= \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
    %    \langle \delta_x(\hat n \chi) \delta_y(\hat n' \chi') \rangle,
    %    \\
    C^{xy}_\ell(\chi,\chi')
    &= \frac{2}{\pi}
\int_0^\infty\ud k k^2 j_\ell(k\chi) j_{\ell}(k\chi')P_{xy}(k).
\end{align}
In reality the power spectrum evolves on the order of a Hubble time.
The angular cross correlations will be very small unless $\chi$ and $\chi'$ are
within a few correlation lengths of one another, roughly a hundred
megaparsecs.  The evolution of the power-spectrum is negligible over these time
differences which leads to a straight-forward way to include the evolution of
the perturbations:
\begin{align}
C^{xy}_\ell(\chi,\chi') 
    \approx \frac{2}{\pi}
    \int_0^\infty\ud k k^2
    j_\ell(k\chi) j_{\ell}(k\chi')
    P_{xy}(k,(\chi + \chi')/2),\nonumber &\\
    |\chi - \chi'| \ll 1/aH.&
\end{align}

@@

There is an advantage to the lack of kernel in the DM-space integral term.
Lensing tomography---the use of sources at multiple redshifts to un-mix
spatial scales---is complicated by the kernel. The kernel's shape depends on
the source redshift, re-weighting the line-of-sight integral and
making tomography inexact.  In contrast the integrand in the DM-space integral
term is completely independent of the limit of integration, meaning
contributions from different parts of the line of sight (and thus different
spatial scales) can be separated exactly. A detailed study of dispersion space
tomography is beyond the scope of this work.  To avoid our results being
sensitive to the hard-to-model small scales, we assume that the contribution to the
integral term from radial distances $\chi'' < 500\,{\rm Mpc}/h$ are well
measured by the correlations with these radial slices. We thus neglect the
information from these slices in our sensitivity measurements and ignore
contributions to the integral term from below this distance. The smallest
spatial scale that contributes the our plotted angular power spectra out to
$\ell=1000$ is thus $k=2\,h/{\rm Mpc}$.

For plotting the DM-space power spectra we use a toy model for the electron and
FRB clustering that assumes both are biased tracers of the dark matter:
\emph{i.e.}~$P_{xy}(k)=b_x b_y P(k)$.  We
calculate the time dependant dark-matter power spectrum using {\tt CAMB} using
the integrated {\tt HALOFIT} to approximate the non-linear evolution. We assume
that the electron bias is $b_e=1$, which should be true on large scales due to
the approximate conservation of free electrons. The bias of FRB sources is
unconstrained by data and we choose it to be $b_f=1.3$. The coefficient
$A(\chi)$ depends on the density of detected FRB sources, $\bar n_f(\chi)$.
We use a simple model where all events above a fixed flux are detected,
the intrinsic luminosity function has the
\citet{1976ApJ...203..297S} form with index $-1$, and events at the cut off of the
luminosity function and at
a radial distance of $\chi=2350\,{\rm Mpc}/h$ are at the flux detection threshold.
This source density is shown in Figure~\ref{f:n_f} with the resulting DM-space
power-spectrum in Figure~\ref{f:terms}.

@@

\begin{figure}
    \includegraphics[scale=0.42]{figures/terms}
    \caption{
        \label{f:terms}
        Terms in the DM-space cross-correlation angular power spectrum.
        All terms are evaluated at $\bar\chi=2000\,{\rm Mpc}/h$ and 
        $\chi - \chi' = 10\,{\rm Mpc}/h$, except the lowest most 
        ``local term'' curve which is for $\chi - \chi'=50\,{\rm Mpc}/h$. 
        While
        the other two terms are highly insensitive to the separation of
        the radial slices, the local term drops rapidly with separation.
    }
\end{figure}

In first measurements of the angular power spectra, uncertainty will be
dominated by the finite number of observed dispersion measures, or shot noise.
The noise power spectrum on shells with of width
$\Delta \chi$ is
\begin{equation}
    C^N_{\ell, ij} = \frac{\delta_{ij}}{\bar n_f(\chi_i)\chi_i^2\Delta\chi}.
\end{equation}
Here, to indicate that we are now working with quantities binned in the radial
direction instead of
continuous functions, we've switched to $C_{\ell,ij}$ over 
$C_\ell(\chi, \chi')$. The uncertainty on the angular
cross-power spectrum is then
\begin{align}
    (\Delta{C^{ss}_{\ell,ij}})^2 = 
    {\frac{1}{\ell(\ell + 1)\Delta \ell f_{\rm sky}}}
        \left[C^N_{\ell,ii} C^N_{\ell,jj} +  (C^N_{\ell,ij})^2\right],
\end{align}
where the second term in brackets only contributes for $i = j$.

There are a number of effects that will complicate the measurement of the
above calculated angular power spectrum by a real survey of dispersion 
measures.  The first effect is
that the model that the free electron density decomposes into a homogeneous
part plus a random field of perturbations is not complete. Firstly, all signals
must pass through the Milky Way's galactic disk, halo, and local environment,
which will contribute to the dispersion measure. However, this is only a
single function of $\hat n$ which can presumably be well
measured. In addition, this function could in principle be well characterized
by independent data such as observations of pulsars.  As such, it should be
possible to subtract the local
contribution, as was done in
\citet{2014arXiv1412.4829D}, and this is unlikely to be a limiting obstacle.

Perhapse, more concerning is the dispersion from the source's environment
as well as the halo in which it resides. These electrons will presumably be
clustered around the source on scales far smaller than DM-space
structure map pixels and thus are not accounted for in the $\delta_e$--$\delta_f$
correlations included in the above formalism.  This will have a mean
contribution, common to all sources and which may be epoch dependant, that will
modify the mean DM--distance relation. This will need to be calibrated in some
why or else fit with nuisance parameters during parameter estimation.  There
will also be a stochastic piece which varies from source to source.  This will
limit the precision at which $\chi_s$ can be measured, thus limiting the
resolution of DM space density maps in the radial direction.

@@

We have proposed using FRB sources as 3D tracers of the large-scale structure,
using dispersion measure to estimate radial distance. It can be seen in
Figure~\ref{f:terms} that the signal is dominated by inhomogeneities in the
electron density along the line of sight inducing mis-estimates of radial
distance and thus apparent clustering. These DM-space distortions could be used
to study the distribution of free electrons in the Universe on large-scales.
There have been other proposals to use FRB events to study large-scale
structure \citep{2014ApJ...780L..33M,2014MNRAS.443L..11D}, however these
schemes have either been two dimensional or required externally measured redshifts
for the sources. Additionally, there have been proposals to use the DM-redshift
relation to probe the background density of electrons and thus the expansion
history of the Universe
\citep{2014ApJ...783L..35D,2014PhRvD..89j7303Z,2014ApJ...788..189G}.

We have
shown that a survey detecting $10^4$ fast radio bursts of cosmological
origin (as might be achieved by the CHIME FRB instrument) could detect the
clustering signal.
Looking forward, DM-space
clustering could become a powerful probe of large-scale structure should
observational factors turn out to be favourable compared to spectroscopic
surveys. The baryon acoustic oscillation feature can be clearly seen in
Figure~\ref{f:terms} in the cross term and both
local-term curves. It may also be possible to extract the BAO feature
from the integral
term using tomography.
This however requires that the contribution to dispersion measure from
the source's environment and host halo be modest compared to the
contribution from the IGM.

>>>>

However, redshift is not the only measure of radial distance.   Standard
candles and standard rulers can be used to estimate distance using brightness
and angular size respectively.  In the same way a \emph{standard ping}, a brief
broadband radio impulse, can be used to estimate radial distance because of the
characteristic $\lambda^2$ dispersion of electromagnetic signals along the
line-of-sight as a function of wavelength.  This technique, routinely done with pulsars in the Galaxy, can be used to define
a \emph{dispersion distance} to cosmological sources.  Intriguingly, the dispersion of fast radio bursts (FRB), a new class
of millisecond duration radio transients first reported by \citet{2007Sci...318..777L}, has been measured with high significance. 
These bursts are the archetype of a standard ping with 
10 reported
\citep{2012MNRAS.425L..71K,2013Sci...341...53T,
2014ApJ...790..101S, 2014ApJ...792...19B, 
2015MNRAS.447..246P,2015ApJ...799L...5R} to date and total rate of bursts
been estimated to be $3.3^{+5}_{-2.5} \times 10^3$
per sky per day \cite{2015arXiv150500834R}. 
Dispersion is characterized by a line-of-site integral of the free electron density known the dispersion measure
(DM).
FRB events are observed to have dispersion measures of order
$\sim\!\!1000 {\rm pc} / {\rm cm}^3$, greatly in excess of
the galactic expectation of order $\sim\!\!100\,{\rm pc} / {\rm cm}^3$ (depending on
galactic latitude)
\cite{2014arXiv1412.4829D}.
They  thus appear to be of extra-galactic origin, with their
dispersion produced either by the
inter-galactic medium (IGM) \citep{2013Sci...341...53T} surrounding galaxies or the
environment of the source
\citep{2014ApJ...785L..26L,2015arXiv150101341P,2015arXiv150505535C}.
If the dispersion is dominated by electrons along the line-of-sight,
then the average dispersion--distance relation can be modelled
\citep{2014ApJ...797...71Z,2004MNRAS.348..999I, 2003ApJ...598L..79I}
and the sources appear to be at
cosmological distances of of order $\sim\!\!{\rm Gpc}$. Such cosmological distances permit
the study of clustering of FRB sources in three dimensions with an expectation that these events trace the
large-scale structure of the Universe on linear scales. Note that although we
frame our discussion in terms of FRBs, other bright radio transients that fluctuate on timescales less than the typical lag time of might also be used as standard pings.

Neither redshift nor dispersion are perfect proxies for radial distance.
Redshifts are systematically biased by the peculiar velocities of the object
relative to the Hubble flow. This leads to additional apparent clustering
in redshift
space which was first calculated by \citet{1987MNRAS.227....1K}
using linear theory.
Similarly, distance estimates from dispersion will be biased by
inhomogeneities in the electron density \citep{2014ApJ...780L..33M},
again leading to additional apparent
clustering of sources in dispersion measure space.  In this letter we calculate
these dispersion-space distortions and consider the detectability of the signal by
upcoming surveys.

%\section{Density perturbations in dispersion measure space}
\section{Clustering in dispersion space}

@@

\emph{Dispersion space} is the three dimensional coordinates,
$\vec x_s$ inferred from the dispersion measure assuming that the electrons are
homogeneous. This only affects the radial coordinate such that $\vec x_s = \hat
n \chi_s$ with $\chi_s$ defined by the equation
\begin{equation}
    {\rm DM}(\chi_s) = \int_0^{\chi_s} \ud\chi' a(\chi')^2 \bar{n}_e(\chi').
\end{equation}

@@

We wish to relate the density of a tracer, $f$, measured in dispersion space to its
density in real space. We follow the derivation in \citet{1987MNRAS.227....1K}
of the
redshift-space distortions. Start by noting that the total number of tracers
in a volume element is the same in both spaces:
\begin{equation}
\label{e:density}
n_{fs}(\vec x_s) \ud^3\vec x_s = n_{f}(\vec x) \ud^3\vec x.
\end{equation}
We split the density into a homogeneous part plus perturbations,
\begin{equation}
    \bar{n}_{fs}(\chi_s)\left[ 1 + \delta_{fs}(\vec x_s)\right] \ud^3\vec x_s
    = \bar{n}_{f}(\chi)\left[ 1 + \delta_f(\vec x)\right] \ud^3\vec x.
\end{equation}
Averaged over the sky, $\langle \chi_s \rangle = \chi$ and thus the background
density should
be the same in dispersion space as in real space,
\begin{equation}
\bar{n}_{fs}(\chi) = \bar{n}_f(\chi).
\end{equation}
Therefore,
\begin{align}
\bar{n}_{fs}(\chi_s) 
    &= \bar{n}_{f}(\chi) + (\chi_s - \chi)\frac{\ud \bar{n}_f}{\ud \chi}\\
    &= \bar{n}_{f}(\chi)
       + \frac{\ud \bar{n}_f}{\ud \chi}\int_0^\chi \ud \chi' \delta_e(\hat n
       \chi').
       \label{e:nfs}
\end{align}

@@

Substituting Equations~\ref{e:nfs}~and~\ref{e:jac} into Equation~\ref{e:density},
we obtain
\begin{equation}
\label{e:delta_s}
    \delta_{fs} = \delta_f - \delta_e
    - \left(\frac{1}{\bar{n}_f}\frac{\ud \bar{n}_f}{\ud \chi}
    + \frac{2}{\chi} \right)
        \int_0^\chi \ud \chi' \delta_e(\hat n \chi').
\end{equation}
In the equation above, the $-\delta_e$ term is most analogous to the Kaiser
redshift-space distortions.  It is a dilution of tracers in dispersion space due to an
excess of electrons between the tracers. However we note that this term is isotropic
in contrast to Kaiser.  This is because any wave vector electron perturbation
causes dispertion space distortions, whereas the radial velocities that cause
redshift-space distortions are only sourced by perturbations with radial wave
vectors.

@@

Large-scale structure is usually studied through its two-point statistics, most
commonly the power spectrum, $P(k)$.
Unlike in redshift-space distortions, Equation~\ref{e:delta_s} does not have a
simple form in harmonic space. The equation's third term couples harmonic modes
and thus the two-point statistics cannot be phrased as a simple power spectrum.
We will instead use $C^{ss}_\ell(\chi,\chi')$, which is
the cross-correlation angular power spectrum of the dispersion-space over density,
on shells at $\chi$ and $\chi'$:
\begin{align}
    &C^{ss}_\ell(\chi, \chi') =
    \int\ud\Omega\ud\Omega' Y_{\ell m}(\hat n) Y^*_{\ell m}(\hat n')
    \langle \delta_{fs}(\hat n \chi) \delta_{fs}(\hat n' \chi')
        \rangle.
\end{align}
The first two terms in Equation~\ref{e:delta_s} are stationary. For stationary,
isotropic, tracers $x$ and $y$ we have 
$\langle \delta_x(\vec k, \chi) \delta_y(\vec k', \chi) \rangle = (2\pi)^3
\delta^3(\vec k - \vec k') P_{xy}(k, \chi)$.  If for a moment we ignore time
dependence, the angular cross-power spectrum of such tracers is
\begin{align}
    %\delta_{\ell\ell'}\delta_{mm'}C^{xy}_\ell(\chi, \chi')
    %&= \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
    %    \langle \delta_x(\hat n \chi) \delta_y(\hat n' \chi') \rangle,
    %    \\
    C^{xy}_\ell(\chi,\chi')
    &= \frac{2}{\pi}
\int_0^\infty\ud k k^2 j_\ell(k\chi) j_{\ell}(k\chi')P_{xy}(k).
\end{align}
In reality the power spectrum evolves on the order of a Hubble time.
The angular cross correlations will be very small unless $\chi$ and $\chi'$ are
within a few correlation lengths of one another, roughly a hundred
megaparsecs.  The evolution of the power-spectrum is negligible over these time
differences which leads to a straight-forward way to include the evolution of
the perturbations:
\begin{align}
C^{xy}_\ell(\chi,\chi') 
    \approx \frac{2}{\pi}
    \int_0^\infty\ud k k^2
    j_\ell(k\chi) j_{\ell}(k\chi')
    P_{xy}(k,(\chi + \chi')/2),\nonumber &\\
    |\chi - \chi'| \ll 1/aH.&
\end{align}

@@

There is an advantage to the lack of kernel in the dispersion-space integral term.
Lensing tomography---the use of sources at multiple redshifts to un-mix
spatial scales---is complicated by the kernel. The kernel's shape depends on
the source redshift, re-weighting the line-of-sight integral and
making tomography inexact.  In contrast the integrand in the dispersion-space integral
term is completely independent of the limit of integration, meaning
contributions from different parts of the line of sight (and thus different
spatial scales) can be separated exactly. A detailed study of dispersion space
tomography is beyond the scope of this work.  To avoid our results being
sensitive to the hard-to-model small scales, we assume that the contribution to the
integral term from radial distances $\chi'' < 500\,{\rm Mpc}/h$ are well
measured by the correlations with these radial slices. We thus neglect the
information from these slices in our sensitivity measurements and ignore
contributions to the integral term from below this distance. The smallest
spatial scale that contributes the our plotted angular power spectra out to
$\ell=1000$ is thus $k=2\,h/{\rm Mpc}$.

For plotting the dispersion-space power spectra we use a toy model for the electron and
FRB clustering that assumes both are biased tracers of the dark matter:
\emph{i.e.}~$P_{xy}(k)=b_x b_y P(k)$.  We
calculate the time dependant dark-matter power spectrum using {\tt CAMB} using
the integrated {\tt HALOFIT} to approximate the non-linear evolution. We assume
that the electron bias is $b_e=1$, which should be true on large scales due to
the approximate conservation of free electrons. The bias of FRB sources is
unconstrained by data and we choose it to be $b_f=1.3$. The coefficient
$A(\chi)$ depends on the density of detected FRB sources, $\bar n_f(\chi)$.
We use a simple model where all events above a fixed flux are detected,
the intrinsic luminosity function has the form from
\citet{1976ApJ...203..297S} with index $-1$, and events at the cut off of the
luminosity function and at
a radial distance of $\chi=2350\,{\rm Mpc}/h$ are at the flux detection threshold.
This source density is shown in Figure~\ref{f:n_f} with the resulting dispersion-space
power-spectrum in Figure~\ref{f:terms}.

@@

\begin{figure}
    \includegraphics[scale=0.42]{figures/terms}
    \caption{
        \label{f:terms}
        Terms in the dispersion-space cross-correlation angular power spectrum.
        All terms are evaluated at $\bar\chi=2000\,{\rm Mpc}/h$ and 
        $\chi - \chi' = 10\,{\rm Mpc}/h$, except the lowest most 
        ``local term'' curve which is for $\chi - \chi'=50\,{\rm Mpc}/h$. 
        While
        the other two terms are highly insensitive to the separation of
        the radial slices, the local term drops rapidly with separation.
    }
\end{figure}

In first measurements of the angular power spectra, uncertainty will be
dominated by the finite number of observed events, or shot noise.
The noise power spectrum on shells with of width
$\Delta \chi$ is
\begin{equation}
    C^N_{\ell, ij} = \frac{\delta_{ij}}{\bar n_f(\chi_i)\chi_i^2\Delta\chi}.
\end{equation}
Here, to indicate that we are now working with quantities binned in the radial
direction instead of
continuous functions, we've switched to $C_{\ell,ij}$ instead of
$C_\ell(\chi, \chi')$. The uncertainty on the angular
cross-power spectrum is then
\begin{align}
    (\Delta{C^{ss}_{\ell,ij}})^2 = 
    {\frac{1}{\ell(\ell + 1)\Delta \ell f_{\rm sky}}}
        \left[C^N_{\ell,ii} C^N_{\ell,jj} +  (C^N_{\ell,ij})^2\right],
\end{align}
where the second term in brackets only contributes for $i = j$.

There are a number of effects that will complicate the measurement of the
above calculated angular power spectrum by a real survey of dispersion 
measures.  The first effect is
that the model that the free electron density decomposes into a homogeneous
part plus a random field of perturbations is not complete. Firstly, all signals
must pass through the Milky Way's galactic disk, halo, and local environment,
which will contribute to the dispersion. However, this is only a
single function of $\hat n$ which can presumably be well
measured. In addition, this function could in principle be well characterized
by independent data such as observations of pulsars.  As such, it should be
possible to subtract the local
contribution, as was done in
\citet{2014arXiv1412.4829D}, and this is unlikely to be a limiting obstacle.

Perhapse, more concerning is the dispersion from the source's environment
as well as the halo in which it resides. These electrons will presumably be
clustered around the source on scales far smaller than dispersion-space
structure map pixels and thus are not accounted for in the $\delta_e$--$\delta_f$
correlations included in the above formalism.  This will have a mean
contribution, common to all sources and which may be epoch dependant, that will
modify the mean dispersion--distance relation. This will need to be calibrated in some
why or else fit with nuisance parameters during parameter estimation.  There
will also be a stochastic piece which varies from source to source.  This will
limit the precision at which $\chi_s$ can be measured, thus limiting the
resolution of dispersion space density maps in the radial direction.

@@

We have proposed using FRB sources as 3D tracers of the large-scale structure,
using dispersion to estimate radial distance. It can be seen in
Figure~\ref{f:terms} that the signal is dominated by inhomogeneities in the
electron density along the line of sight inducing mis-estimates of radial
distance and thus apparent clustering. These dispersion-space distortions could be used
to study the distribution of free electrons in the Universe on large-scales.
There have been other proposals to use FRB events to study large-scale
structure \citep{2014ApJ...780L..33M,2014MNRAS.443L..11D}, however these
schemes have either been two dimensional or required externally measured redshifts
for the sources. Additionally, there have been proposals to use the dispersion-redshift
relation to probe the background density of electrons and thus the expansion
history of the Universe
\citep{2014ApJ...783L..35D,2014PhRvD..89j7303Z,2014ApJ...788..189G}.

We have
shown that a survey detecting $10^4$ fast radio bursts of cosmological
origin (as might be achieved by the CHIME FRB instrument) could detect the
clustering signal.
Looking forward, dispersion-space
clustering could become a powerful probe of large-scale structure should
observational factors turn out to be favourable compared to spectroscopic
surveys. The baryon acoustic oscillation feature can be clearly seen in
Figure~\ref{f:terms} in the cross term and both
local-term curves. It may also be possible to extract the BAO feature
from the integral
term using tomography.
This however requires that the contribution to dispersion from
the source's environment and host halo be small compared to the
contribution from the IGM.
